Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-24 Paper/conc.tex
..

In this work, we focused on the uncertainty associated with the
rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism and its
impact on ignition delay predictions. The mechanism involves 19 different
reactions and in each case, the reaction rate depends upon the choice of a
pre-exponent and an activation energy. Hence, in theory, the evolution of the
chemical system depends upon 38 inputs. 
\alennote{Maybe say something about how we went from 38 to 33?} 
%Conventional means for uncertainty quantification such as those involving
%surrogate model construction as well as sensitivity analysis would thus be
%computationally challenging.
To facilitate efficient uncertainty analysis, we focused our efforts on
reducing the dimensionality of the problem by identifying important directions
in the parameter space such that the variability in the model output is
predominantly captured along these directions. This was done through active
subspace construction.  Additionally, we demonstrated that the activity scores,
computed using the components of the dominant eigenvectors provide an efficient
means for approximating derivative based global sensitivity measures (DGSMs).
Furthermore, we established generalized mathematical linkages between the
different global sensitivity measures: activity scores, DGSMs, and total Sobol'
index which could be exploited to reduce computational effort associated with
global sensitivity analysis.

@@

Additionally, the surrogates were assessed by comparing probability
distributions for the ignition delay,
constructed using surrogate predictions and model evaluations at the same set
of 10$^4$ random samples. The PDFs based on model evaluations and the
1-dimensional surrogate from the perturbation-based strategy were found to be
almost identical. However, the PDF based on 1-dimensional surrogate from the
regression-based strategy although captured the modal estimate with reasonable
accuracy, was observed to underestimate the uncertainty in the ignition delay.
Mean estimates of the ignition delay using model evaluations and predictions
using the two surrogates at 10$^4$ samples were found to be in agreement.
However, the standard deviation, computed using predictions from
the surrogate based on the regression-based strategy was found to be accurate
only upto the first significant digit. Moreover, it was observed that the
activity scores, computed using the regression-based strategy failed to capture
the sensitivity towards one of the most important parameters, $E_{a,15}$.  We attribute
this discrepancy to two potential sources of numerical errors: (1)
Regression-based approximation of the gradient of the model output, and (2)
Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
error incurred by the linear regression fit in this case. Hence, while the
regression-based approach 
reduces computational effort, the present investigations clearly reveal
potential pitfalls associated with this approach. 
%
%More specifically, using
%the regression-based approach, 
%\begin{itemize} 
%\item The desired statistics (e.g. mean and variance) of the
%QoI might not be approximated to a desired level of accuracy.  
%\item The
%approach might fail to capture the sensitivity  of the QoI towards all inputs.
%\end{itemize} 
Therefore, in general perturbation-based approaches are preferred. The
regression-based approach can be explored in situations involving intensive
simulations where the perturbation-based methods are intractable, and the goal
is obtaining rough estimates of the statistics of the QoI as opposed to a more
detailed analysis such as those involving parametric sensitivity.  We also
mention that alternate regression based approaches such as ones based on
computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{ConstantineDoostan17}. The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

>>>>

In this work, we focused on the uncertainty associated with the
rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism and its
impact on ignition delay predictions. The mechanism involves 19 different
reactions and in each case, the reaction rate depends upon the choice of a
pre-exponent and an activation energy. Hence, in theory, the evolution of the
chemical system depends upon 38 inputs. In this work, we have considered 
epistemic uncertainty in all pre-exponents and activation energies with non-zero
nominal values. Hence, the input space dimensionality is 33 instead of 38.  
%Conventional means for uncertainty quantification such as those involving
%surrogate model construction as well as sensitivity analysis would thus be
%computationally challenging.
To facilitate efficient uncertainty analysis, we focused our efforts on
reducing the dimensionality of the problem by identifying important directions
in the parameter space such that the variability in the model output is
predominantly captured along these directions. This was done through active
subspace construction.  Additionally, we demonstrated that the activity scores,
computed using the components of the dominant eigenvectors provide an efficient
means for approximating derivative based global sensitivity measures (DGSMs).
Furthermore, we established generalized mathematical linkages between the
different global sensitivity measures: activity scores, DGSMs, and total Sobol'
index which could be exploited to reduce computational effort associated with
global sensitivity analysis.

@@

Additionally, the surrogates were assessed by comparing probability
distributions for the ignition delay,
constructed using surrogate predictions and model evaluations at the same set
of 10$^4$ random samples. The PDFs based on model evaluations and the
1-dimensional surrogate from the perturbation-based strategy were found to be
almost identical. However, the PDF based on 1-dimensional surrogate from the
regression-based strategy although captured the modal estimate with reasonable
accuracy, was observed to underestimate the uncertainty in the ignition delay.
Mean estimates of the ignition delay using model evaluations and predictions
using the two surrogates at 10$^4$ samples were found to be in agreement.
However, the standard deviation, computed using predictions from
the surrogate based on the regression-based strategy was found to be accurate
only upto the first significant digit. Moreover, it was observed that the
activity scores, computed using the regression-based strategy failed to capture
the sensitivity towards one of the most important parameters, $E_{a,15}$.  We attribute
this discrepancy to two potential sources of numerical errors: (1)
Regression-based approximation of the gradient of the model output, and (2)
Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
error incurred by the linear regression fit in this case. Hence, while the
regression-based approach 
reduces computational effort, the present investigations clearly reveal
potential pitfalls associated with this approach. 
%
%More specifically, using
%the regression-based approach, 
%\begin{itemize} 
%\item The desired statistics (e.g. mean and variance) of the
%QoI might not be approximated to a desired level of accuracy.  
%\item The
%approach might fail to capture the sensitivity  of the QoI towards all inputs.
%\end{itemize} 
Therefore, in general perturbation-based approaches are preferred. The
regression-based approach can be explored in situations involving intensive
simulations where the perturbation-based methods are intractable, and the goal
is obtaining rough estimates of the statistics of the QoI as opposed to a more
detailed analysis such as those involving parametric sensitivity.  We also
mention that alternate regression based approaches such as ones based on
computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{Constantine:2017a}. The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

### 2018-09-24 Paper/intro.tex
..

In this work, we implement two strategies for computing the active subspace for the
H$_2$/O$_2$ mechanism, and refer to them as \textit{perturbation-based} and 
\textit{regression-based} to highlight the difference in the approach for estimating
the gradient of the model output as discussed.  
Other than the difference in gradient estimation, the two strategies follow the same
sequence of steps provided in~\cite{Constantine:2015}. However, 
in order to reduce computational effort, the two strategies are implemented in an
iterative manner in this work as discussed later in Section~\ref{sec:method}.

An alternate approach to dimension reduction involves computing the global
sensitivity measures associated with the uncertain inputs of a model. Depending
upon the estimates of the sensitivity measures, only the important inputs are
varied for the purpose of uncertainty quantification (UQ). Sobol' indices are
commonly used as global sensitivity measures~\cite{Sobol:2001}. They are
used to quantify the relative contributions of the uncertain inputs to the variance
in the output, either individually, or in combination with other inputs. 
Multiple efforts have focused on efficient computation of the Sobol' 
indices~\cite{Sudret:2008,Plischke:2013,Tissot:2015,Li:2016} including the 
derivative-based global sensitivity measures~(DGSMs), developed to
compute approximate upper bounds for the Sobol' indices with much fewer
computations~\cite{Sobol:2009, Lamboni:2013}. It was noted
in~\cite{Diaz:2016,Constantine:2017} that DGSMs can be approximated by
exploiting their links with active subspaces. This led to the definition of the 
so-called \emph{activity scores}. In Section~\ref{sub:gsa}, we build on these
ideas to provide a complete analysis of links between Sobol indices, DGSMs, and
activity scores for functions of independent random inputs whose distribution
law belongs to the class of Boltzmann probability measures. 
It is worth mentioning that computing global sensitivity measures provides 
important information about a model that go beyond dimension reduction. By 
identifying parameters with significant impact on the model output, we can assess
regimes of validity of the model formulation, and gain critical insight into the
underlying physics in many cases.

The main contributions of this paper are as follows: 
\begin{itemize}
\item 
Active subspace
discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 33
uncertain inputs: The methodology presented in this work
resulted in a 1-dimensional active subspace, indicating immense potential
for computational savings. The presented
analysis can also guide practitioners in other problems of chemical kinetics on using the
method of active subspaces to achieve efficiency in uncertainty propagation.  
\item Comprehensive numerical investigation of the perturbation-based and the regression-based approaches: 
We investigate the suitability of both approaches
for computing active subspaces, surrogate modeling, and approximating global
sensitivity measures through a comprehensive set of numerical experiments. Our
results reveal insight into the merits of the methods as well as
their shortcomings.  
\item Analysis of the links between  
global sensitivity measures: 
By connecting the recent theoretical advances in variance-based and
derivative-based global sensitivity analysis to active subspaces, we provide a complete analysis
of the links between total Sobol' indices, DGSMs, and activity scores for a broad
class of probability distributions. Our analysis is concluded by a result quantifying
approximation errors incurred due to fixing unimportant parameters, deemed so by 
computing their activity scores.

@@

This article is organized as follows. In section~\ref{sub:ac}, a brief
theoretical background on the active subspace methodology is provided. In
section~\ref{sub:gsa}, it is shown that the activity scores provide a
reasonable approximation to the DGSMs especially in a high-dimensional setting.
Additionally, a relationship between the three global sensitivity measures,
namely, the activity scores, DGSMs, and the total Sobol' indices is
established. In section~\ref{sec:method}, a systematic framework for computing
the active subspace is provided. 
Numerical results based on the perturbation-based approach are compared with those
obtained using the regression-based approach.  For our initial investigation aimed
as assessing the feasibility and suitability of the two strategies, the active 
subspace is computed for a 19-dimensional 
H$_2$/O$_2$ reaction kinetics problem whereby only the $A_i$'s are
considered as uncertain. In this case, the two approaches are observed to yield
consistent results. Motivated by these results, we further implement the two
strategies to compute the active subspace for a 33-dimensional H$_2$/O$_2$
reaction kinetics problem in section~\ref{sec:app}. Once again, the two
strategies reveal the existence of a 1-dimensional active subspace. However,
it is observed that although the regression-based strategy results in a 
reasonable approximation of the mean, mode, and the uncertainty in the ignition
delay, it fails to capture the global sensitivity of all the uncertain inputs. 
Hence, a trade-off between efficiency and accuracy between the two approaches is
apparent in this case.  Finally, a summary and discussion based on our findings is included in
section~\ref{sec:conc}.

>>>>

In this work, we adopt an iterative strategy to reduce the computational effort
associated with active subspace computation. Moreover, we explore two approaches
for estimating the gradient of the ignition delay with respect to the uncertain
rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism. As mentioned above, the
first approach uses finite difference to estimate the gradient and is regarded as
the perturbation-based approach, whereas, the second approach involves a linear
regression-fit to the available set of model evaluations and is regarded as
the regression-based approach.

An alternate strategy to dimension reduction involves computing the global
sensitivity measures associated with the uncertain inputs of a model. Depending
upon the estimates of the sensitivity measures, only the important inputs are
varied for the purpose of uncertainty quantification (UQ). Sobol' indices are
commonly used as global sensitivity measures~\cite{Sobol:2001}. They are
used to quantify the relative contributions of the uncertain inputs to the variance
in the output, either individually, or in combination with other inputs. 
Multiple efforts have focused on efficient computation of the Sobol' 
indices~\cite{Sudret:2008,Plischke:2013,Tissot:2015,Li:2016} including the 
derivative-based global sensitivity measures~(DGSMs), developed to
compute approximate upper bounds for the Sobol' indices with much fewer
computations~\cite{Sobol:2009, Lamboni:2013}. It was noted
in~\cite{Diaz:2016,Constantine:2017} that DGSMs can be approximated by
exploiting their links with active subspaces. This led to the definition of the 
so-called \emph{activity scores}. In Section~\ref{sub:gsa}, we build on these
ideas to provide a complete analysis of links between Sobol indices, DGSMs, and
activity scores for functions of independent random inputs whose distribution
law belongs to the class of Boltzmann probability measures. 
It is worth mentioning that computing global sensitivity measures provides 
important information about a model that go beyond dimension reduction. By 
identifying parameters with significant impact on the model output, we can assess
regimes of validity of the model formulation, and gain critical insight into the
underlying physics in many cases.

The main contributions of this paper are as follows: 
\begin{itemize}
\item 
Active subspace
discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 33
uncertain inputs: The methodology presented in this work
resulted in a 1-dimensional active subspace, indicating immense potential
for computational savings. The presented
analysis can also guide practitioners in other problems of chemical kinetics on using the
method of active subspaces to achieve efficiency in uncertainty propagation.  
\item Comprehensive numerical investigation of the perturbation-based and the regression-based approaches: 
We investigate the suitability of both approaches
for estimating the gradient of ignition delay in the H$_2$/O$_2$ mechanism.
Specifically, we compare resulting
active subspaces, surrogate models, and the ability to approximate global
sensitivity measures through a comprehensive set of numerical experiments. Our
results reveal insight into the merits of the methods as well as
their shortcomings.  
\item Analysis of the links between  
global sensitivity measures: 
By connecting the recent theoretical advances in variance-based and
derivative-based global sensitivity analysis to active subspaces, we provide a complete analysis
of the links between total Sobol' indices, DGSMs, and activity scores for a broad
class of probability distributions. Our analysis is concluded by a result quantifying
approximation errors incurred due to fixing unimportant parameters, deemed so by 
computing their activity scores.

@@

This article is organized as follows. In section~\ref{sub:ac}, a brief
theoretical background on the active subspace methodology is provided. In
section~\ref{sub:gsa}, it is shown that the activity scores provide a
reasonable approximation to the DGSMs especially in a high-dimensional setting.
Additionally, a relationship between the three global sensitivity measures,
namely, the activity scores, DGSMs, and the total Sobol' indices is
established. In section~\ref{sec:method}, a systematic framework for computing
the active subspace is provided. 
Numerical results based on the perturbation-based approach are compared with those
obtained using the regression-based approach.  The active 
subspace is initially computed for a 19-dimensional 
H$_2$/O$_2$ reaction kinetics problem wherein only the $A_i$'s are
considered as uncertain. In this case, the two approaches are observed to yield
consistent results. We further compute the active  for a 33-dimensional H$_2$/O$_2$
reaction kinetics problem in section~\ref{sec:app}. Once again, 
active subspace is found to be 1-dimensional. However, in this case,
it is observed that although the regression-based estimation of the gradient results in a 
reasonable approximation of the mean and mode, the variability in the ignition delay
due to uncertain rate-controlling parameters is under-estimated. Additionally,  
the sensitivity towards one of the most important parameter is not captured.
Hence, a trade-off between efficiency and accuracy between the two approaches is
apparent in the higher-dimensional problem. 
Finally, a summary and discussion based on our findings is included in
section~\ref{sec:conc}.

### 2018-09-24 Paper/method.tex
..

In this section, we present an efficient framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Specifically, we employed two strategies: 
a perturbation-based strategy that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression-based approach 
that involves a local linear approximation of
the model output to approximate the gradient. 
The two approaches are based on Algorithms~1.1 and 1.2 in~\cite{Constantine:2015} and
essentially differ in the way the gradient is estimated. They were implemented in an
iterative manner to ensure computational efficiency by avoiding 
unnecessary model evaluations once a converged active subspace
was achieved.

@@

As discussed earlier, gradient estimation using finite differences in the 
perturbation-based strategy requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, gradient estimation
using the regression-based approach involves a series of linear regression
fits to subsets of available evaluations as discussed 
in~\cite{Constantine:2015}. Hence, the computational effort is  
reduced by a factor $(d+1)$ when using the regression-based approach.
The specific sequence of steps
for computing the active subspace is discussed as follows.

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
at an initial set of $n_0$ samples denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition
of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
$\hat{\mat{W}}_1$ and the set of corresponding eigenvalues, 
$\hat{\mat{\Lambda}}_1$. Note that $\hat{\mat{W}}_1$ is obtained by
partitioning the eigenspace around $\lambda_j$ such that the ratio of
subsequent eigenvalues,
$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)>\mathcal{O}(10^2)$.
 At each subsequent iteration, model evaluations are
generated at a new set of $n_k$ samples. The new set of gradient evaluations
are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
by its eigenvalue decomposition. The relative change in the norm of the
difference in squared value of individual components of the dominant
eigenvectors between subsequent iterations is evaluated. The process is
terminated and the resulting eigenspace is considered to have converged once
the maximum relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$
($j$ is used as an index for the eigenvectors),
is smaller than a given tolerance, $\tau$.  A regression fit to
$G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and
quantify the uncertainty in the model output. Moreover, the components of the
dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
which provide an insight into the relative importance of the uncertain inputs.
Note that the index, $r$, corresponds to the number of dominant eigenvectors in
$\hat{\mat{W}}_1$. The sequence of steps as discussed are outlined
in Algorithm~\ref{alg:grad}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain $A_i$'s. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
\lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
$N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly three orders of magnitude larger than $\lambda_2$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

>>>>

In this section, we present an efficient framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Two approaches are explored for estimating the gradient of ignition delay with
respect to uncertain $A_i$'s: 
a perturbation-based approach that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression-based approach 
that involves a linear regression fit to the available set of
model evaluations in order to approximate the gradient. 
The active subspace is computed in an iterative manner to avoid
unnecessary model evaluations once converged is established.

@@

As discussed earlier, gradient estimation using finite differences 
requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, gradient estimation
using the regression-based approach involves a series of linear regression
fits to subsets of available evaluations as discussed 
in~\cite{Constantine:2015} (see Algorithm 1.2). Hence, the computational effort is  
reduced by a factor $(d+1)$ when using the regression-based approach.
The specific sequence of steps
for computing the active subspace is discussed as follows.

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
at an initial set of $n_0$ samples (generated using Monte Carlo sampling)
denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition
of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
$\hat{\mat{W}}_1$ and the set of corresponding eigenvalues, 
$\hat{\mat{\Lambda}}_1$. Note that $\hat{\mat{W}}_1$ is obtained by
partitioning the eigenspace around $\lambda_j$ such that the ratio of
subsequent eigenvalues,
$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)>\mathcal{O}(10^2)$.
 At each subsequent iteration, model evaluations are
generated at a new set of $n_k$ samples. The new set of gradient evaluations
are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
by its eigenvalue decomposition. The relative change in the norm of the
difference in squared value of individual components of the dominant
eigenvectors between subsequent iterations is evaluated. The process is
terminated and the resulting eigenspace is considered to have converged once
the maximum relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$
($j$ is used as an index for the eigenvectors),
is smaller than a given tolerance, $\tau$.  A regression fit to
$G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and
quantify the uncertainty in the model output. Moreover, the components of the
dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
which provide an insight into the relative importance of the uncertain inputs.
Note that the index, $r$, corresponds to the number of dominant eigenvectors in
$\hat{\mat{W}}_1$. The sequence of steps as discussed are outlined
in Algorithm~\ref{alg:grad}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain $A_i$'s. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum ($\lambda_i/\lambda_0,~i = 1,\ldots,19$)
with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum ($\lambda_i/\lambda_0$)
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly three orders of magnitude larger than $\lambda_2$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.
